---
title: "scBind_SC_ST"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = dirname(getwd()))
```

# Intro

BiCCA

<https://htmlpreview.github.io/?https://github.com/KChen-lab/bindSC/blob/master/vignettes/SC_ST/SC_ST.html>

Seurat Integration

<https://satijalab.org/seurat/archive/v3.2/spatial_vignette.html>

<https://satijalab.org/seurat/archive/v3.0/integration.html>

<https://satijalab.org/seurat/articles/integration_introduction.html>

Deconvolution

<https://nbisweden.github.io/workshop-scRNAseq/labs/compiled/seurat/seurat_07_spatial.html#Deconvolution>

Seurat SCT normalization

<https://satijalab.org/seurat/articles/sctransform_vignette.html>

# Article

<https://www.biorxiv.org/content/10.1101/2020.12.11.422014v1.full>

# Notes

1.  regress mitochondria content does not seems to change the data transfer results much
2.  Normalization by NormalizeData (scaling+log) or SCT does not seems to change the data transfer results much
3.  Seurat TransferData: really important to use features from the integration method.
4.  The RNA-RNA gene expression imputation is generally much better using Seurat than bind SC.
5.  Although bindSC can impute all gene expression, it is unclear if these imputation can be trusted
6.  For t-cycIf + scRNAseq, using log transformed, but not scaled data seems to work much better

# Library

```{r echo=FALSE, message=FALSE}
library(Seurat)
library(sctransform)
library(bindSC)
```

# Parameters

```{r}
mito = 0 # mitochondria regression
n_components = 3
SCT = 0 # normalization SCT =0 is NormalizeData
```

# Functions ---

## Function - co-embedding double plot

```{r}
# assumption: in the umap_3D matrix, cat1 rows, followed by cat2 rows
double_plot <- function(umap_3D, cat1, cat2, source1, source2) {
  cat1_label = levels(source1)[1]
  cat2_label = levels(source2)[1]
  
  umap_plt  <- data.frame("UMAP1"=umap_3D[,1],
                          "UMAP2"=umap_3D[,2],
                          "cell_annotation" = c(cat1, cat2),
                          "data_source" = c(rep(cat1_label, length(source1)), 
                                            rep(cat2_label, length(source2))))
  

  xlim <- c(min(umap_plt$UMAP1), max(umap_plt$UMAP1))
  ylim <- c(min(umap_plt$UMAP2), max(umap_plt$UMAP2))
  
  p11 <- UMAP_plot(meta = umap_plt[umap_plt$data_source==cat1_label,], 
                   mylabel = paletteDiscrete(umap_plt$cell_annotation), 
                   alpha=1, color = "cell_annotation", xlim = xlim, ylim = ylim) + ggtitle(cat1_label)
  
  p12 <- UMAP_plot(meta = umap_plt[umap_plt$data_source==cat2_label,], 
                   mylabel=paletteDiscrete(umap_plt$cell_annotation),
                  alpha = 1, color = "cell_annotation", xlim = xlim, ylim = ylim) + ggtitle(cat2_label)

  ggarrange(p11,p12, ncol=2)
}
```

## Function - save co-embedding cohort data (umap3D, data source, cell annotation)

```{r}
# assumption: in the umap_3D matrix, cat1 rows, followed by cat2 rows
save_coembedding <- function(umap_3D, cat1, cat2, source1, source2, prefix="") {
  cat1_label = levels(source1)[1]
  cat2_label = levels(source2)[1]
  
  write.table (umap_3D, 
               paste0("umap3D_coembedding", suffix, ".tsv"), 
               quote=FALSE, sep="\t",  col.names = NA)

  write.table (data.frame("cell_annotation" = c(paste0(cat1,"_", source1), 
                                                paste0(cat2,"_", source2)), 
                          "data_source" = c(rep(cat1_label,length(cat1)), 
                                            rep(cat2_label,length(cat2))),
                          row.names = rownames(umap_3D)),
             paste0(prefix, "cell_annotation_coembedding", ".tsv"),
             quote=FALSE, sep="\t",  col.names = NA)
}
```

## Function - confusion matrix plots

```{r}
confusion_matrix <- function(rowMeta, colMeta, suffix) {
  library(corrplot)
  table_with_data<- data.frame("rowMeta" = rowMeta, "colMeta" = colMeta)
  
  generatePlot<- function() {
    confus_mat <- as.matrix(table(table_with_data[,c("rowMeta","colMeta")]))
    for(i in seq(1, nrow(confus_mat), 1)){
      confus_mat[i,] <- confus_mat[i,]/sum(confus_mat[i,])
    }
    corrplot(as.matrix(confus_mat),is.corr = FALSE)
    
    confus_mat <- as.matrix(table(table_with_data[,c("colMeta","rowMeta")]))
    for(i in seq(1, nrow(confus_mat), 1)){
      confus_mat[i,] <- confus_mat[i,]/sum(confus_mat[i,])
    }
    corrplot(as.matrix(confus_mat),is.corr = FALSE)  
  }
  
  pdf(file=paste0("confus_mat",suffix,".pdf"))
  generatePlot()
  dev.off()
  
  generatePlot()
}
```

## Function - impuZ_2 imputation

```{r}
# dt1 bindSC co-embedding - X
# dt2 bindSC co-embedding

# imputing features (any continuous feature) in dt1/X for samples in dt2
impuZ_2 <- function(X, dt1, dt2){
    out <- X%*%dt1%*%t(dt2)
    return(out)
}
```

## Function - GeneImp2 imputation

```{r}
# dt1 bindSC co-embedding: on RNA cells 
# dt2 bindSC co-embedding: on Merfish cells 
# X : gene expression matrix on RNA cells, or any type of continuous data from RNA (dt1) 
# k:  neighbor size

# imputing features (any continuous feature) in dt1/X for samples in dt2
GeneImp2 <- function(dt1 = NULL, X = NULL, dt2 = NULL, k=NULL){
    dis<-cdist(as.matrix(dt2), as.matrix(dt1))
    # using k-neighbour to update coembeddings 
    v <- rep(0, nrow(dt2))
    for(i in seq(1, nrow(dis),1)){
          b<-sort(dis[i,])
          v[i]= b[k]
          d<-rep(0, length(b))
          d[which(dis[i,]<=b[k])]<-1
          dis[i,]<-d/k
    }
    U <- dis%*%t(X)
    return(list(t(U),v))
}

```

## Function - export to Xena

```{r}
toXena <- function(seuratObject, writeExp = TRUE) {
  # metadata
  data = seuratObject@meta.data
  write.table(data.frame("ID"=rownames(data), data), file = "meta.tsv", sep='\t', quote=FALSE, row.names=F)
  
  # expression matrix
  if (writeExp) {
    data = seuratObject[[seuratObject@active.assay]]@data
    write.table(data.frame("ID"=rownames(data), data), file = "expression.tsv", sep='\t', quote=FALSE, row.names=F)
  }
  
  # umap reduction
  if (exists("umap", where= seuratObject@reductions)) {
    data = seuratObject@reductions$umap@cell.embeddings
    write.table(data.frame("ID"=rownames(data), data), file = "umap3D.tsv", sep='\t', quote=FALSE, row.names=F)
  }
}
```

# Load scRNA-seq (Allen brain) and pp

MT%, filtering, scTransform normalization, [PCA, neighbor, UMAP, clustering] - don't really need this since the data has class, subclass and cluster information in the metadata already One cortical scRNA-seq data composed of \~14,000 adult mouse primary visual cortex (VISp) from the Allen Institute was collected. This dataset was generated using the SMART-Seq2 protocol. The data source for download is from Seurat (<https://www.dropbox.com/s/cuowvm4vrf65pvq/allen_cortex.rds?dl=1>).

```{r}
getwd()
SC <- readRDS(file="input_data/allen_cortex.rds")

library(tidyverse)
mouse_mito = as_tibble(read.csv("../Mouse.MitoCarta3.0.csv", header = TRUE))
mouse_mito = mouse_mito %>% select(c(Symbol, MitoCarta2.0_Score))

mito.genes = as.character(mouse_mito$Symbol)
mito.genes = mito.genes[mito.genes %in% rownames(SC[['RNA']]@data)]
percent.mito = Matrix::colSums(SC[['RNA']]@data[mito.genes,]) / Matrix::colSums(SC[['RNA']]@data) * 100
SC = AddMetaData(object = SC, metadata = percent.mito, col.name = "percent.mito")
Idents(SC) <- "SC"

p1<- VlnPlot(SC, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)
ggarrange(p1)

SC <- subset(SC, subset =  nFeature_RNA < 12500 & 
               nFeature_RNA > 2500 & 
               nCount_RNA < 3.5e+06 &
               percent.mito < 20)
SC$orig.ident <- SC$subclass

# normalization
if (SCT) {  
  if (mito){
    SC <- SCTransform(SC, vars.to.regress = "percent.mito")
  } else {
    SC <- SCTransform(SC)
  }
  DefaultAssay(SC) <- "SCT"  
} else {
  DefaultAssay(SC) <- "RNA"
  SC<-NormalizeData(SC)
  SC<-FindVariableFeatures(SC, selection.method = "vst", nfeatures = 2000)
  if (mito){
    SC<-ScaleData(SC, vars.to.regress = c("percent.mito")) %>% 
      RunPCA()
  } else {
    SC<-ScaleData(SC) %>% 
      RunPCA()
  }
}

unique(SC$subclass)
y.clst <- SC$subclass
```

# Load scRNA-seq (liver met) and pp

```{r}
datafile = "/Users/jzhu/Documents/JingData/WORK/Cancer/Single_Cell/16-080L/scRNAseq/filtered_feature_bc_matrix.h5"
clusterfile ="/Users/jzhu/Documents/JingData/WORK/Cancer/Single_Cell/16-080L/scRNAseq/cluster.tsv"
```

```{r}
SC.rna <- Read10X_h5(datafile)
cluster <- read.csv(file = clusterfile, sep = '\t', header = TRUE)

SC <- CreateSeuratObject(counts = SC.rna,  meta.data = cluster)

SC$orig.ident<- SC$louvain
```

```{r}
p1<- VlnPlot(SC, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2)
ggarrange(p1)
p1
```

```{r}
SC <- subset(SC, subset =  nFeature_RNA < 6000 & 
               nFeature_RNA > 200 & 
               nCount_RNA < 30000 )
```

```{r}
SC<-NormalizeData(SC, scale.factor =1e+6)
SC<-FindVariableFeatures(SC, selection.method = "vst", nfeatures = 5000)
SC<-ScaleData(SC)
```

```{r}
Idents(SC) <- "scRNA-seq"
y.clst <- SC$louvain
```

# Load scRNA-seq (melanoma) and pp

data: <https://singlecell.broadinstitute.org/single_cell/study/SCP109/melanoma-immunotherapy-resistance> The datafile is in the unit of log2(1+TPM/10 ), already subset to only the non-tumor cells, and already normalized against the total value per cell, and scale to the level of 1 million 1e+6

```{r}
# non malignant cells only
datafile = "/Users/jzhu/Documents/JingData/WORK/Cancer/Single_Cell/melanoma_Regiv2018/SCP109/integrationData/tumors.nonmal_tpm.tsv"
metafile ="/Users/jzhu/Documents/JingData/WORK/Cancer/Single_Cell/melanoma_Regiv2018/SCP109/integrationData/tumors.nonmal_anno.txt"

# malignant and non malignant cells 
# datafile = "/Users/jzhu/Documents/JingData/WORK/Cancer/Single\ Cell/melanoma_Regiv2018/SCP109/integrationData/tumors.all_tpm.txt"

# metafile ="/Users/jzhu/Documents/JingData/WORK/Cancer/Single\ Cell/melanoma_Regiv2018/SCP109/integrationData/cell.type.all.tsv"

SC.rna <- as.sparse(read.csv(file = datafile, sep = "\t", header = TRUE, row.names = 1))
# revert back from  log2(1+TPM/10 )
SC.rna  = (2^SC.rna-1)*10

meta <- read.csv(file = metafile, sep = '\t', header = TRUE, row.names = 1)

SC <- CreateSeuratObject(counts = SC.rna,  meta.data = meta)
SC$orig.ident<- SC$cell.type

# since need to integrate with t-CycIF data, which has only 43 markders. Need to get a new smaller matrix with just these 43 markders with the matching antibody name

DefaultAssay(SC) <- "RNA"
SC<-NormalizeData(SC, scale.factor =1e+6)
SC<-FindVariableFeatures(SC, selection.method = "vst", nfeatures = 2000)

SC<-ScaleData(SC) %>% 
  RunPCA() %>%
  FindNeighbors() %>%
  FindClusters()    

SC <- RunUMAP(SC, n.components = 3L,  dims = 1:10)
DimPlot(SC, reduction = "umap", dims = c(1, 2))

Idents(SC) <- "SC"
#y.clst <- SC$cell.type
y.clst <- SC$seurat_clusters

```

```{r}
toXena(SC, writeExp = FALSE)
```

```{r}
confusion_matrix(SC$orig.ident, SC$seurat_clusters, "_cluster")
```

# Load abEstimate

```{r}
# non malignant cells only
datafile = "/Users/jzhu/Documents/JingData/WORK/Cancer/Single\ Cell/melanoma_Regiv2018/SCP109/integrationData/tumors.nonmal_AbEstimate.tsv"
metafile ="/Users/jzhu/Documents/JingData/WORK/Cancer/Single\ Cell/melanoma_Regiv2018/SCP109/integrationData/tumors.nonmal_anno.txt"

# maliganant and non malignant cells
# datafile = "/Users/jzhu/Documents/JingData/WORK/Cancer/Single\ Cell/melanoma_Regiv2018/SCP109/integrationData/tumors.all_AbEstimate.tsv"
# metafile ="/Users/jzhu/Documents/JingData/WORK/Cancer/Single\ Cell/melanoma_Regiv2018/SCP109/integrationData/cell.type.all.tsv"


Z.rna <- t(as.sparse(read.csv(file = datafile, sep = "\t", header = TRUE, row.names = 1)))
meta <- read.csv(file = metafile, sep = '\t', header = TRUE, row.names = 1)

Z <- CreateSeuratObject(counts = Z.rna,  meta.data = meta)
Idents(Z) <- "estimateAB"
Z$orig.ident<- Z$cell.type

DefaultAssay(Z) <- "RNA"
Z[['RNA']]@data <- log1p(Z[['RNA']]@counts)

Z<-FindVariableFeatures(Z, selection.method = "vst")

Z<-ScaleData(Z) %>% 
    RunPCA()
```

# Load ST (vizgen mouse coronal) and pp

Vizgen MERFISH Mouse Brain Receptor Map S2R1 count data <https://info.vizgen.com/mouse-brain-map?submissionGuid=556602ae-7d8e-400d-998d-f01939580490> It takes 30min to load the tsv file, it seems much much longer than scanpy loading tsv file, maybe it will be easier to save the data as .annadata then convert it to .seurat type of data.

```{r}
# this code generate a large dgCMatrix, with is what the BiCCA needs, note that BiCCA does not need seurat object
datafile = "input_data/datasets_mouse_brain_map_BrainReceptorShowcase_Slice2_Replicate1_cell_by_gene_S2R1.tsv"

ST.rna <- as.sparse(read.csv(file = datafile, sep = "\t", header = TRUE, row.names = 1))
colnames(ST.rna) = substring(colnames(ST.rna),2)

meta <- read.csv(file = 'input_data/vizgenMeta.tsv', sep = '\t', header = TRUE, colClasses = c("X"="character"))
row.names(meta) = meta$X
ST <- CreateSeuratObject(counts = ST.rna,  meta.data = meta)
Idents(ST) <- "ST"
ST$orig.ident<- ST$louvain
  
# there are only two mitochondria genes. no regress out mito signature
ST <- NormalizeData(ST)
x.clst <- ST$louvain
```

# Load ST (xenium tsv) and pp

```{r}
datafile = "/Users/jzhu/Documents/JingData/WORK/Cancer/Single_Cell/16-080L/xenium/exprMatrix.tsv"
metafile ="/Users/jzhu/Documents/JingData/WORK/Cancer/Single_Cell/16-080L/xenium/cells.tsv"
clusterfile = "/Users/jzhu/Documents/JingData/WORK/Cancer/Single_Cell/16-080L/xenium/cluster.tsv"
```

```{r}
ST.rna <- as.sparse(read.csv(file = datafile, sep = "\t", header = TRUE, row.names = 1))
meta <- read.csv(file = metafile, sep = '\t', header = TRUE)
cluster <- read.csv(file = clusterfile, sep = '\t', header = TRUE)
meta<-merge(meta, cluster, by = "xenium_id", all.x = TRUE)

ST <- CreateSeuratObject(counts = ST.rna,  meta.data = meta, row.names = 1)

ST$orig.ident<- ST$louvain
```

```{r}
cells_to_keep <- (!is.na(ST@meta.data$louvain)) & (ST@meta.data$nFeature_RNA >1)
ST <- ST[, cells_to_keep]
```

```{r}
expr_matrix <- GetAssayData(ST, slot = "data")
gene_sd <- apply(expr_matrix, 1, sd)
sd_threshold <- 0.15
genes_to_keep <- names(gene_sd)[gene_sd > sd_threshold]
ST[["RNA"]]@var.features <-genes_to_keep
```

```{r}
ST <- NormalizeData(ST, scale.factor = 1e+6)
ST<-ScaleData(ST)

Idents(ST) <- "xenium"
x.clst <- ST$louvain
```

# Load ST (t-CycIF melanoma) and pp

data: <https://www.tissue-atlas.org/atlas-datasets/nirmal-maliga-vallius-2021>. MEL08-1-1

```{r}
datafile = "/Users/jzhu/Documents/JingData/WORK/Cancer/Single_Cell/Melanoma_sorger2022/MEL08-1-1.raw.tsv"
metafile ="/Users/jzhu/Documents/JingData/WORK/Cancer/Single_Cell/Melanoma_sorger2022/celltype.tsv"
xycorfile = "/Users/jzhu/Documents/JingData/WORK/Cancer/Single_Cell/Melanoma_sorger2022/MEL08-1-1.csv"
clusterfile = "/Users/jzhu/Documents/JingData/WORK/Cancer/Single_Cell/Melanoma_sorger2022/cluster.tsv"

ST.rna <- t(as.sparse(read.csv(file = datafile, sep = "\t", header = TRUE, row.names = 1)))
meta <- read.csv(file = metafile, sep = '\t', header = TRUE)

ST <- CreateSeuratObject(counts = ST.rna,  meta.data = meta, row.names = 1)

new = read.csv(file= xycorfile)
ST@meta.data["Y_centroid"] = new$Y_centroid
new = read.csv(file= clusterfile, sep='\t')

ST@meta.data["leiden"] = new$leiden
ST@meta.data["louvain"] = new$louvain

ST$orig.ident<- ST$phenotype_brief
```

```{r}
p1<- VlnPlot(ST, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2)
ggarrange(p1)

ST <- subset(ST, subset = nCount_RNA < 1.8e+05)
```

```{r}
# remove the bottom sample, only keep the top sample
#ST = subset(x=ST, subset = Y_centroid < 7500)

# remove Melanocytes (tumor) data and Unknown data
Idents(ST) <- ST$phenotype_brief
ST = subset(x = ST, idents = c("Melanocytes", "Unknown"), invert = TRUE)

ST <- NormalizeData(ST, scale.factor = 1e+6)
ST<-ScaleData(ST)

Idents(ST) <- "t-CycIF"
#x.clst <- ST$phenotype_brief
x.clst <- ST$louvain
```

# Load ST (Visium mouse sagittal anterior1) pp

direct from Seurat

```{r}
library(SeuratData)
if (!("stxBrain.SeuratData" %in% rownames(InstalledData()))) {
    InstallData("stxBrain")
}
ST <- LoadData("stxBrain", type = "anterior1")

mito.genes = as.character(mouse_mito$Symbol)
mito.genes = mito.genes[mito.genes %in% rownames(ST)]
percent.mito = Matrix::colSums(ST[mito.genes,]) / Matrix::colSums(ST) * 100
ST = AddMetaData(object = ST, metadata = percent.mito, col.name = "percent.mito")
Idents(ST) <- "ST"

p1 <- VlnPlot(ST, features = c("nFeature_Spatial", "nCount_Spatial", "percent.mito"), ncol = 3)
ggarrange(p1)

if (SCT)  {  
  if (mito) {
    ST <- SCTransform(ST, vars.to.regress = "percent.mito", assay = "Spatial") %>%
        RunPCA() %>%
        FindNeighbors(dims = 1:30) %>%
        RunUMAP(dims = 1:30) %>%
        FindClusters()
  } else {
    ST <- SCTransform(ST, assay = "Spatial") %>%
        RunPCA() %>%
        FindNeighbors(dims = 1:30) %>%
        RunUMAP(dims = 1:30) %>%
        FindClusters()
  }
  DefaultAssay(SC) <- "SCT"  
} else {
  DefaultAssay(ST) <- "Spatial"
  ST<-NormalizeData(ST)
  ST<-FindVariableFeatures(ST, selection.method = "vst", nfeatures = 2000)
  if (mito) {
    ST<-ScaleData(ST, vars.to.regress = c("percent.mito"))
  } else {
    ST<-ScaleData(ST)
  }
  ST <- RunPCA(ST) %>%
        FindNeighbors(dims = 1:30) %>%
        RunUMAP(dims = 1:30) %>%
        FindClusters()
}

ST$orig.ident<- ST$seurat_clusters
x.clst <- ST$seurat_clusters
```

# BiCCA method ---

## Prepare for the input for bindSC alignment, since X and Z needs the same dimension, a subset of ST and SC overlapping genes

X: ST\
Z0:ST feature, SC cell\
Y: SC

### ST, SC

```{r}
if (SCT) {
  ST_genes<-ST[["SCT"]]@var.features
  SC_genes<-SC[["SCT"]]@var.features
  gene.use <- intersect(ST_genes, SC_genes)
  x <- ST[['SCT']]@scale.data[gene.use,]
  z0 <- SC[['SCT']]@scale.data[gene.use,]
  y<- SC[['SCT']]@scale.data
} else {
  ST_genes<-ST[["RNA"]]@var.features
  SC_genes<-SC[["RNA"]]@var.features
  gene.use <- intersect(ST_genes, SC_genes)
  x <- ST[['RNA']]@scale.data[gene.use,]
  z0 <- SC[['RNA']]@scale.data[gene.use,]
  y<- SC[['RNA']]@scale.data
  
  #gene.use <- integration.features #Seurat integration method features
  #x <- ST[['Spatial']]@data[gene.use,]
  #z0 <- SC[['RNA']]@data[gene.use,]
  #y<- SC[['RNA']]@data
}

#y.clst <- SC@meta.data$subclass
```

### t-CycIF, SC

```{r}
x <- ST[['RNA']]@scale.data

z0<- Z[['RNA']]@scale.data

y <- SC[['RNA']]@scale.data

#y <-SC[['RNA']]@data[rownames(SC[['RNA']]@data) %in% rownames(SC[['RNA']]@scale.data), ]  

```

## run BiCCA without dimensionality reduction

```{r}
res <- BiCCA( X = x ,
             Y = y, 
             Z0 = z0, 
             X.clst = x.clst,
             Y.clst = y.clst,
             alpha = 0.5, 
             lambda = 0.5,
             K = 15,
             temp.path  = "out",
             num.iteration = 50,
             tolerance = 0.01,
             save = TRUE,
             parameter.optimize = FALSE,
             block.size = 0)
```

## CO-embedding double plot

use python method - fast

```{r}
library(reticulate)
umap_import <- import(module = "umap", delay_load = TRUE)
umap_import$pkg_resources$get_distribution("umap-learn")
# example https://github.com/satijalab/seurat/blob/master/R/dimensional_reduction.R#L1305

umap.args = list(densmap = FALSE, n_components = n_components) 
umap <- do.call(what = umap_import$UMAP, args = umap.args)
umap_3D <- umap$fit_transform(rbind(res$u, res$r))

if (n_components == 3){
  colnames(umap_3D) =c("UMAP1","UMAP2","UMAP3")  
}
if (n_components == 2){
  colnames(umap_3D) =c("UMAP1","UMAP2")  
}

double_plot(umap_3D, x.clst, y.clst, Idents(ST), Idents(SC))

rownames(umap_3D) =c(dimnames(res$u)[[1]],dimnames(res$r)[[1]])
```

## Save the joint embedding, data sources, orig cell annotation

```{r}
prefix = "bindSC_"
save_coembedding(umap_3D, ST$orig.ident, SC$orig.ident, Idents(ST), Idents(SC), prefix)

```

#################################################################################################### 

## function - label transfer function using SVM

```{r}
# Input:
#   dt1 FROM latent representation 
#   dt2 TO latent representation in the same space
#   X.clst FROM label
# Return:
#   $prob_max TO probability for the most likely label
#   $prob_full TO full probability matrix 

label_transfer_full_prob_SVM <- function(dt1 = NULL, X.clst = NULL, dt2 = NULL, k=3) {
    library(bindSC)
    dis<-cdist(as.matrix(dt2), as.matrix(dt1))
    # using k-neighbour to update coembeddings 
    #k <- 3
    for(i in seq(1, nrow(dis),1)){
          b<-sort(dis[i,])
          d<-rep(0, length(b))
          d[which(dis[i,]<=b[k])]<-1
          dis[i,]<-d/k
    }
    U <- dis%*%dt1
        

    model <- svm(dt1, as.factor(X.clst), probability=TRUE)
    pred <- predict(model, U, probability=TRUE)
    out <- attr(pred, "probabilities")
   
    celltype <- rep("unknown", nrow(out))
    prob_max <- rep(0, nrow(out))
    data_point <- rep(k, nrow(out))
    names <- colnames(out)
    for(i in seq(1,nrow(out),1)){
          tp <- out[i,]
          pos <- which(tp==max(tp),arr.ind = TRUE)
          celltype[i] <- names[pos]
          prob_max[i] <- tp[pos]
    }
    prob <- data.frame("predict_celltype"=celltype, "Prob_max"=prob_max, "data_point"=data_point)
    rownames(prob) <- rownames(dt2)
    rownames(out) <- rownames(dt2)
    #colnames(out) <- paste0(colnames(out),suffix)
    ret <- list()
    ret$prob_max = prob
    ret$prob_full = out
    return(ret)
}
```

## categorical data label transfer from ST to SC

```{r}
library(e1071)
k=3
suffix="_bindSC"
ret <- label_transfer_full_prob_SVM (dt1=res$u, X.clst=x.clst, dt2=res$r, k=k)
cell_type_predict_SC <- ret$prob_max
cell_type_prob_full_SC <- ret$prob_full
head(cell_type_predict_SC)
```

## Save the predicted cell type for SC data

```{r}
dataTable = data.frame (
  "cell_type_predict"=paste0(cell_type_predict_SC$predict_celltype, "_predict"), 
  "Prob_max"= cell_type_predict_SC$Prob_max)
colnames(dataTable) = c(paste0("cell_type_pred", suffix), paste0("Prob_max", suffix))

write.table(cbind("ID"=rownames(cell_type_predict_SC), dataTable, cell_type_prob_full_SC),
            paste0("cell_type_predict", suffix, ".tsv"),
            quote=FALSE, sep="\t", row.names = F)
```

## categorical data label transfer from SC to ST using SVM

```{r}
library(e1071)
k= 3
prefix= "bindSC_"
ret <- label_transfer_full_prob_SVM (dt1=res$r, X.clst=y.clst, dt2=res$u, k=k)
cell_type_predict_ST <- ret$prob_max
cell_type_prob_full_ST <- ret$prob_full
head(cell_type_predict_ST)
head(cell_type_prob_full_ST)
```

## Save the predicted cell type for ST data

```{r}
dataTable = data.frame (
  "cell_type_predict"= cell_type_predict_ST$predict_celltype, 
  "Prob_max"= cell_type_predict_ST$Prob_max)
colnames(dataTable) = c(paste0(prefix, "cell_type_pred"), paste0(prefix, "Prob_max"))

write.table(cbind("ID"=rownames(cell_type_predict_ST), dataTable, cell_type_prob_full_ST),
            paste0(prefix, "labelTransfer", ".tsv"),
            quote=FALSE, sep="\t", row.names = F)

# dataTable = data.frame (
#   "celltype_predict_annotate"=c(paste0(cell_type_predict_ST$predict_celltype,"_predict"), paste0(y.clst, "_annotated")),
#   "Prob_max"= c(cell_type_predict_ST$Prob_max, rep("",ncol(SC))),
#   "data_point"= c(cell_type_predict_ST$data_point, rep("",ncol(SC))))
# colnames(dataTable) = c(paste0("celltype_predict_annotate",suffix),
#                         "Prob_max","data_point")
# write.table (dataTable,
#              paste0("cell_type_predict_coembedding", suffix,".tsv"),
#              quote=FALSE, sep="\t", col.names = NA,
#              row.names = c(row.names(cell_type_predict_ST),
#                            dimnames(res$r)[[1]]))

```

## Confusion matrix

```{r}
confusion_matrix(x.clst, cell_type_predict_ST$predict_celltype, suffix)
```

# Seurat Integration method ---

## Integration by CCA

<https://satijalab.org/seurat/archive/v3.0/integration.html>

```{r}
if (exists("Z")){
  datasetList = c(Z,ST)
} else {
  datasetList = c(SC,ST)
}
integration.features= SelectIntegrationFeatures(object.list = datasetList)

if (SCT) {
  integration.list = PrepSCTIntegration(object.list = datasetList, anchor.features = integration.features)
  
  integration.anchors <- FindIntegrationAnchors(object.list = integration.list, 
                                                anchor.features = integration.features, 
                                                normalization.method = "SCT")
  seurat_integrated <- IntegrateData(anchorset = integration.anchors, normalization.method = "SCT")
} else {
  integration.anchors <- FindIntegrationAnchors(object.list = datasetList, 
                                                anchor.features = integration.features)
  seurat_integrated <- IntegrateData(anchorset = integration.anchors)
  seurat_integrated <- ScaleData(seurat_integrated)
}

seurat_integrated <- RunPCA(seurat_integrated)
seurat_integrated <- RunUMAP(seurat_integrated, dims = 1:20, umap.method = "umap-learn", n.components = n_components)
```

## Co-embedding double plot CCA

```{r}
umap_3D = as.matrix(seurat_integrated$umap@cell.embeddings)
double_plot(umap_3D, SC$orig.ident, ST$orig.ident, Idents(SC), Idents(ST))
```

## Integration by reciprocal PCA (RPCA)

<https://satijalab.org/seurat/articles/integration_rpca.html>

```{r}
datasetList = c(SC,ST)
integration.features= SelectIntegrationFeatures(object.list = datasetList)
integration.list = PrepSCTIntegration(object.list = datasetList, anchor.features = integration.features)
integration.list <- lapply(X = integration.list, FUN = RunPCA, features = integration.features)

integration_rpca.anchors <- FindIntegrationAnchors(object.list = integration.list, 
                                         normalization.method = "SCT",
                                         anchor.features = integration.features, 
                                         dims = 1:30, reduction = "rpca", k.anchor = 20)
seurat_integrated_rpca <- IntegrateData(anchorset = integration_rpca.anchors, 
                                     normalization.method = "SCT", dims = 1:30)

seurat_integrated_rpca <- RunPCA(seurat_integrated_rpca)
seurat_integrated_rpca <- RunUMAP(seurat_integrated_rpca, 
                                  reduction = "pca", dims = 1:30, 
                                  umap.method = "umap-learn",
                                  n.components = n_components)
```

## Co-embedding double plot RPCA

```{r}
umap_3D = as.matrix(seurat_integrated_rpca$umap@cell.embeddings)
double_plot(umap_3D, SC$orig.ident, ST$orig.ident, Idents(SC), Idents(ST))
```

## save joint embedding, data source, and original annotation

```{r}
suffix = "_Seurat_Integration"
save_coembedding(umap_3D, SC$orig.ident, ST$orig.ident, Idents(SC), Idents(ST), suffix)
```

write.table(umap_plt_object)

## Categorical Data Transfer

<https://satijalab.org/seurat/archive/v3.2/spatial_vignette.html>

```{r}
suffix = "_Seurat_Integration"

if (exists("Z")){
  query = Z
} else {
  query = ST
}

if (SCT) {
  anchors <- FindTransferAnchors(reference = SC, query = query, normalization.method = "SCT", features= integration.features)
} else {
  anchors <- FindTransferAnchors(reference = SC, query = query, normalization.method = "LogNormalize", features= integration.features)
}
predictions <- TransferData(anchorset = anchors, refdata = SC$orig.ident, k.weight=7)
```

## save predictions

```{r}
write.table(predictions, paste0("cell_type_predict", suffix, ".tsv"), quote = FALSE, sep = '\t', col.names = NA)
```

## confusion matrix

```{r}
confusion_matrix(ST$orig.ident, predictions$predicted.id, suffix)
```

# BiCCA imputation

## Impute x (ST) feature for y (SC) samples

```{r}
#imputation of featurs in the entire x (ST) matrix for y (SC) samples, can be used to any x feature
input_x = x
Z_est <- res$Z_est
Z_impu <- impuZ_2(X=input_x, dt1 = res$u, dt2=res$r)
Z_impu2 <- GeneImp2(res$u, x, res$r, 3)
colnames(Z_impu) <- colnames(z0)
colnames(Z_impu2) <- colnames(z0)
```

```{r}
# gene in bicca
gene_ST = 'pERK'
gene_SC= 'MAPK3'

orig_y = SC[[SC@active.assay]]@data[gene_SC,] 
orig_x= ST[[ST@active.assay]]@data[gene_ST,]

table = data.frame("Z_impu"=Z_impu[gene_ST,], 
                   "Z_impu2" = Z_impu2[gene_ST,],
                   "y"=orig_y, 
                   "Z_est" = Z_est[gene_ST,])

s1 <-cor.test(table$Z_impu2, table$y)
s2 <-cor.test(table$Z_impu, table$y)
s3 <-cor.test(table$Z_est, table$y)
s4 <-cor.test(table$Z_impu, table$Z_impu2)

p1= ggplot(data=table, aes(x = Z_impu2, y = y)) + geom_point(alpha = 0.1) + 
  annotation_custom (grobTree(textGrob(paste0(gene_ST,"\n",
                                              "r:", signif(s1$estimate), "\n",
                                              "p:", signif(s1$p.value,3)))))

p2= ggplot(data=table, aes(x = Z_impu, y = y)) + geom_bin2d() + theme_classic() +
  scale_fill_gradient(low="lightblue1",high="darkblue",trans="log10") + 
  annotation_custom (grobTree(textGrob(paste0(gene_ST,"\n",
                                              "r:", signif(s2$estimate), "\n",
                                              "p:", signif(s2$p.value,3)))))

p3= ggplot(data=table, aes(x = Z_est, y = y)) + geom_bin2d() + theme_classic() +
  scale_fill_gradient(low="lightblue1",high="darkblue",trans="log10") + 
  annotation_custom (grobTree(textGrob(paste0(gene_ST,"\n",
                                              "r:", signif(s3$estimate), "\n",
                                              "p:", signif(s3$p.value,3)))))

p4= ggplot(data=table, aes(x = Z_impu2, y = Z_impu)) + geom_bin2d() + theme_classic() +
  scale_fill_gradient(low="lightblue1",high="darkblue",trans="log10") + 
  annotation_custom (grobTree(textGrob(paste0(gene_ST,"\n",
                                              "r:", signif(s4$estimate), "\n",
                                              "p:", signif(s4$p.value,3)))))

ggarrange(p1, p2, p3, p4)
```

## Impute y (SC) feature for x (ST) samples

```{r}
#imputation of featurs in the y (SC) matrix for x (ST) samples, can be used to any y feature
input_y = t(data.matrix(SC@meta.data %>% select(starts_with("OE."))))
Z_impu <- impuZ_2(X=input_y, dt1 = res$r, dt2=res$u)
r <- GeneImp2(res$r, input_y, res$u, 3)
Z_impu2 <-r[[1]]
Z_impu2_threshold <- data.frame(r[[2]])

colnames(Z_impu) <- colnames(x)
colnames(Z_impu2) <- colnames(x)
colnames(Z_impu2_threshold) <- "impu_threshold"
```

```{r}
# save imputed y feature for x samples
dataTable = t(Z_impu)
write.table(cbind("ID"=rownames(dataTable), dataTable),
            "imputed.tsv",
            quote=FALSE, sep="\t", row.names = F)

dataTable = cbind(t(Z_impu2), 
write.table(cbind("ID"=rownames(dataTable), dataTable),
            "imputed2.tsv",
            quote=FALSE, sep="\t", row.names = F)
```

# Seurat integration of gene expression

```{r}
sub_SC = subset(seurat_integrated, idents="SC")
sub_ST = subset(seurat_integrated, idents="ST")

# gene in integration feature # integration.features
gene = 'Ttr'
table = data.frame("integrated"= sub_ST[['integrated']]@scale.data[gene,], 
                   "origin"=ST[['SCT']]@data[gene,])

s <-cor.test(table$integrated, table$origin)
p1= ggplot(data=table, aes(x = origin, y = integrated)) + geom_point(alpha = 0.1)
p2= ggplot(data=table, aes(x = origin, y = integrated)) + geom_bin2d() + theme_classic() +
  scale_fill_gradient(low="lightblue1",high="darkblue",trans="log10") + 
  annotation_custom (grobTree(textGrob(paste0(gene,"\n",
                                              "r:", signif(s$estimate), "\n",
                                              "p:", signif(s$p.value,3)))))
ggarrange(p1, p2)
```
